---
title: "Getting started with the app"
description: >
  A screenshot walkthrough of the six-step DataGangeR workflow, from picking an
  objective to exporting a synthetic data bundle.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
fig <- function(name) knitr::include_graphics(name)  # frames live beside this Rmd
```

The DataGangeR Shiny app guides you from a real dataset to a shareable synthetic
double in six steps. Launch it with:

```r
library(dataganger)
run_app()
```

Every screen mirrors a plain function call, so anything you do here you can also
script from the R console (see *Use it from R* in the
[README](https://github.com/lennon-li/dataganger#use-it-from-r)).

## 1. Objective

Start by telling DataGangeR what the synthetic data is *for*. Instead of
separate Fidelity / Privacy / Anonymity meters, the app now shows one
**Protection** meter: **Demo / Teaching = 5 bars**, **Development and
prototyping = 3 bars**, and **Internal Analytics = 1 bar**. **Development** is
the default objective.

When you click an objective, the detail panel explains it along the same
dimensions every time: **Use when**, **Exact values**, **Distributions**,
**Relationships**, **Identifiers**, and **Sensitive & rare values**. That keeps
the trade-off readable without making you interpret multiple competing meters.

```{r, echo = FALSE}
fig("step-1-objective.png")
```

## 2. Upload

Drop in a CSV, Excel, or SAS file — or load one of the built-in sample datasets
to explore the workflow without any real data. The right-hand panel previews
your data live as it loads.

```{r, echo = FALSE}
fig("step-2-upload.png")
```

## 3. Configure

This is where you, the human, make the decisions. Nothing is generated until you
have reviewed them. The screen is built around privacy-first column
classification, plus the synthesis settings described below.

```{r, echo = FALSE}
fig("step-3-configure.png")
```

**Per-column decisions.** Each row now asks two intrinsic questions:

- **Does this point to a person?**
  - **No** — blood pressure, lab value, score, price, outcome
  - **Only combined with other columns** — age, sex, ZIP / postcode, birth
    date, job title
  - **Yes, directly** — name, email, phone, address, Social Security number,
    record / account number
- **Is it sensitive?**
  - **No** or **Yes** — mark **Yes** for values like diagnosis, income,
    religion, or anything that would cause harm if linked back to a person

You do **not** choose the action separately in the normal flow. DataGangeR
derives it and shows the result live in **What we'll do**:

|                      | sensitive = No | sensitive = Yes |
|---|---|---|
| identifies = none | Synthesized; distribution kept, exact values not. | Synthesized; protected from linkage. |
| identifies = combination | Coarsened & grouped (k-anonymity), then synthesized. | Coarsened & grouped + protected from linkage. |
| identifies = direct | **Removed** from the output. | **Removed** from the output. |

Direct identifiers always drop. Everything non-direct is synthesized. Every
column still needs an explicit answer to the first question before you can
generate.

**Synthesis settings.** The objective you picked presets safe defaults, but you
can override any of these:

| Setting | What you are deciding |
|---|---|
| **Row count (n)** | How many synthetic rows to generate. |
| **Seed** | Fixes the random draw so the same settings reproduce the exact same data. |
| **Engine** | *Auto* (from objective), *Internal* (per-column, fast, ignores relationships), or *synthpop* (relationship-aware, higher fidelity). |
| **Column name handling** | Keep original names, replace with generic names, or anonymize names and keep the mapping in the data dictionary. |
| **Rare category threshold** | Values seen fewer than this many times count as rare, so they can be merged or suppressed. |
| **Coarsen dates** | Round dates (e.g. to month/year) so an exact date can't single out an individual. |
| **Merge rare categories** | Combine infrequent values into an "other" group to reduce re-identification risk. |
| **Free-text handling** | How free-text columns are treated (usually set by your objective). |
| **Preserve missing values** | How closely to reproduce the original pattern of missing (`NA`) values. |

Defaults are safe for the objective you picked — leave them unless you have a
reason to change them. Each control in the app also carries a one-line
explanation, and every decision is recorded in the exported spec so the result
is reproducible (see [`synth_spec()`](../reference/synth_spec.html) for the
programmatic equivalents).

## 4. Generate

Generate the synthetic double. The summary reports row and column counts, the
engine used, and the elapsed time; the right panel switches to preview the
synthetic records.

```{r, echo = FALSE}
fig("step-4-generate.png")
```

## 5. Compare

Inspect how faithfully the synthetic data mirrors the original. Pick any
variable to see overlaid distributions (green is your original data, magenta is
synthetic), a Q–Q plot, and a side-by-side statistics table with the delta
between them.

```{r, echo = FALSE}
fig("step-5-compare.png")
```

## 6. Export

The in-app download is the **compact** bundle: a small, shareable package built
around six files:

- `synthetic_data.csv`
- `README.md`
- `load_data.R`
- `analysis.qmd`
- `data_dictionary.csv`
- `manifest.json`

If a rendered comparison is included, you also get `comparison_report.html`.

In the compact bundle, `README.md` is consolidated: it includes the privacy
summary and a **For AI assistants** section, so there is no separate
`ai-readme.md` or `privacy_report.txt`. Those standalone files stay in the
**full** CLI / agent bundle (`compact = FALSE` or `make_agent_bundle()`), along
with the extra diagnostics used in agent workflows.

```{r, echo = FALSE}
fig("step-6-export.png")
```

> **Remember:** synthetic data reduces direct disclosure risk; it does not
> replace a formal privacy assessment. Review the comparison and privacy
> warnings before sharing any output externally.

## Reporting problems and feedback

If something feels off, you can open a pre-filled GitHub issue straight from R:

```r
report_issue("The export summary was confusing", context = "Shiny app")
```

The Shiny app also includes a **Report a problem** button in the sidebar. Both
paths open your browser with environment details filled in, but they do not
send anything automatically - you stay in control of what gets submitted.
