---
title: "Getting started with the app"
description: >
  A screenshot walkthrough of the six-step DataGangeR workflow, from picking an
  objective to exporting a synthetic data bundle.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
fig <- function(name) knitr::include_graphics(name)  # frames live beside this Rmd
```

The DataGangeR Shiny app guides you from a real dataset to a shareable synthetic
double in six steps. Launch it with:

```r
library(dataganger)
run_app()
```

Every screen mirrors a plain function call, so anything you do here you can also
script from the R console (see *Use it from R* in the
[README](https://github.com/lennon-li/dataganger#use-it-from-r)).

## 1. Objective

Start by telling DataGangeR what the synthetic data is *for*. The objective —
**Demo / Teaching**, **Development and prototyping**, or **Internal Analytics** —
presets sensible defaults for fidelity, privacy hardening, and coarsening across
the rest of the workflow.

```{r, echo = FALSE}
fig("step-1-objective.png")
```

## 2. Upload

Drop in a CSV, Excel, or SAS file — or load one of the built-in sample datasets
to explore the workflow without any real data. The right-hand panel previews
your data live as it loads.

```{r, echo = FALSE}
fig("step-2-upload.png")
```

## 3. Configure

This is where you, the human, make the decisions. Nothing is generated until you
have reviewed them. There are two kinds of choice on this screen.

```{r, echo = FALSE}
fig("step-3-configure.png")
```

**Per-column decisions.** For every column you set:

- **Disclosure role** — how sensitive the column is for re-identification:
  *direct identifier* (e.g. name, ID), *quasi-identifier* (combines with others
  to single someone out, e.g. age + postcode), *sensitive* (private attribute),
  or *none*. Roles drive privacy hardening and k-anonymity. Every column must
  carry an explicit role before you can generate — uncertain columns are left
  unselected on purpose, so you make the call rather than the tool guessing.
- **Action** — what happens to the column: *synthesize* (replace with synthetic
  values), *pass through* (copy as-is — only safe for non-identifying columns),
  or *drop* (exclude from the output entirely).

**Synthesis settings.** The objective you picked presets safe defaults, but you
can override any of these:

| Setting | What you are deciding |
|---|---|
| **Row count (n)** | How many synthetic rows to generate. |
| **Seed** | Fixes the random draw so the same settings reproduce the exact same data. |
| **Engine** | *Auto* (from objective), *Internal* (per-column, fast, ignores relationships), or *synthpop* (relationship-aware, higher fidelity). |
| **Column name handling** | Keep original names, replace with generic names, or anonymize names and keep the mapping in the data dictionary. |
| **Rare category threshold** | Values seen fewer than this many times count as rare, so they can be merged or suppressed. |
| **Coarsen dates** | Round dates (e.g. to month/year) so an exact date can't single out an individual. |
| **Merge rare categories** | Combine infrequent values into an "other" group to reduce re-identification risk. |
| **Free-text handling** | How free-text columns are treated (usually set by your objective). |
| **Preserve missing values** | How closely to reproduce the original pattern of missing (`NA`) values. |

Defaults are safe for the objective you picked — leave them unless you have a
reason to change them. Each control in the app also carries a one-line
explanation, and every decision is recorded in the exported spec so the result
is reproducible (see [`synth_spec()`](../reference/synth_spec.html) for the
programmatic equivalents).

## 4. Generate

Generate the synthetic double. The summary reports row and column counts, the
engine used, and the elapsed time; the right panel switches to preview the
synthetic records.

```{r, echo = FALSE}
fig("step-4-generate.png")
```

## 5. Compare

Inspect how faithfully the synthetic data mirrors the original. Pick any
variable to see overlaid distributions (green is your original data, magenta is
synthetic), a Q–Q plot, and a side-by-side statistics table with the delta
between them.

```{r, echo = FALSE}
fig("step-5-compare.png")
```

## 6. Export

Download a single bundle containing the synthetic data, the synthesis spec, a
manifest, a comparison report, diagnostic files, and (optionally) a standalone
data dictionary - ready to share with collaborators or AI programming tools.

```{r, echo = FALSE}
fig("step-6-export.png")
```

> **Remember:** synthetic data reduces direct disclosure risk; it does not
> replace a formal privacy assessment. Review the comparison and privacy
> warnings before sharing any output externally.

## Reporting problems and feedback

If something feels off, you can open a pre-filled GitHub issue straight from R:

```r
report_issue("The export summary was confusing", context = "Shiny app")
```

The Shiny app also includes a **Report a problem** button in the sidebar. Both
paths open your browser with environment details filled in, but they do not
send anything automatically - you stay in control of what gets submitted.
