---
title: "Getting started with the app"
description: >
  A screenshot walkthrough of the six-step DataGangeR workflow, from picking an
  objective to exporting a synthetic data bundle.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
fig <- function(name) knitr::include_graphics(name)  # frames live beside this Rmd
```

The DataGangeR Shiny app guides you from a real dataset to a shareable synthetic
double in six steps. Launch it with:

```r
library(dataganger)
run_app()
```

Every screen mirrors a plain function call, so anything you do here you can also
script from the R console (see *Use it from R* in the
[README](https://github.com/lennon-li/dataganger#use-it-from-r)).

## 1. Objective

Start by telling DataGangeR what the synthetic data is *for*. The objective —
**Demo / Teaching**, **Development and prototyping**, or **Internal Analytics** —
presets sensible defaults for fidelity, privacy hardening, and coarsening across
the rest of the workflow.

```{r, echo = FALSE}
fig("step-1-objective.png")
```

## 2. Upload

Drop in a CSV, Excel, or SAS file — or load one of the built-in sample datasets
to explore the workflow without any real data. The right-hand panel previews
your data live as it loads.

```{r, echo = FALSE}
fig("step-2-upload.png")
```

## 3. Configure

Review each column's **disclosure role** (direct identifier, quasi-identifier,
sensitive, or none) and adjust synthesis settings only if you need to. Defaults
are safe for the objective you picked. Every column must carry an explicit role
before you can generate — uncertain columns are left unselected on purpose.

```{r, echo = FALSE}
fig("step-3-configure.png")
```

## 4. Generate

Generate the synthetic double. The summary reports row and column counts, the
engine used, and the elapsed time; the right panel switches to preview the
synthetic records.

```{r, echo = FALSE}
fig("step-4-generate.png")
```

## 5. Compare

Inspect how faithfully the synthetic data mirrors the original. Pick any
variable to see overlaid distributions (green is your original data, magenta is
synthetic), a Q–Q plot, and a side-by-side statistics table with the delta
between them.

```{r, echo = FALSE}
fig("step-5-compare.png")
```

## 6. Export

Download a single bundle containing the synthetic data, a data dictionary, the
synthesis spec, a manifest, a comparison report, and diagnostic files — ready to
share with collaborators or AI programming tools.

```{r, echo = FALSE}
fig("step-6-export.png")
```

> **Remember:** synthetic data reduces direct disclosure risk; it does not
> replace a formal privacy assessment. Review the comparison and privacy
> warnings before sharing any output externally.
