---
title: "DataGangeR: original vs synthetic comparison"
format:
  html:
    toc: true
    df-print: paged
params:
  original_path: "path/to/your/original-data.csv"
  synthetic_path: "synthetic_data.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 4
)
library(dataganger)
library(ggplot2)
```

{%purpose_line}

This report shows the R-only workflow for reproducing this synthetic data and,
when your original file is available, comparing it against the exported
synthetic dataset.

## Reproduce this synthetic data

```{r reproduction, eval=FALSE}
{%reproduction_script}
```

## Read both datasets

```{r read}
synthetic_dir <- dirname(params$synthetic_path)
if (identical(synthetic_dir, "")) {
  synthetic_dir <- "."
}
source(file.path(synthetic_dir, "load_data.R"), local = TRUE)
synthetic <- load_dataganger_bundle(synthetic_dir)

has_original <- file.exists(params$original_path)
if (has_original) {
  original <- read_input(params$original_path)
  shared <- intersect(names(original), names(synthetic))
  original <- original[, shared, drop = FALSE]
  synthetic <- synthetic[, shared, drop = FALSE]

  combined <- rbind(
    cbind(original, .source = "original"),
    cbind(synthetic, .source = "synthetic")
  )
  combined$.source <- factor(combined$.source, levels = c("original", "synthetic"))
} else {
  original <- NULL
  shared <- names(synthetic)
  combined <- NULL
  message(
    "Original data not found at '",
    params$original_path,
    "'. Set params$original_path to your real input file to run the comparisons."
  )
}

# DataGangeR convention: green = original, magenta = synthetic.
dg_cols <- c(original = "#3F8F4F", synthetic = "#D81B8C")
```

## Summary statistics

```{r summary}
if (is.null(original)) {
  message("Skipping original-vs-synthetic summary until params$original_path exists.")
  summary(synthetic)
} else {
  summary(original)
  summary(synthetic)
}
```

## Distribution comparisons

Green is the original data; magenta is the synthetic data.

```{r numeric-plots}
if (is.null(original)) {
  message("Skipping numeric comparison plots until params$original_path exists.")
} else {
  numeric_cols <- {%numeric_cols}
  numeric_cols <- intersect(numeric_cols, shared)
  for (v in numeric_cols) {
    print(
      ggplot(combined, aes(x = .data[[v]], fill = .source)) +
        geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
        scale_fill_manual(values = dg_cols) +
        labs(title = v, x = v, fill = NULL) +
        theme_minimal()
    )
  }
}
```

```{r categorical-plots}
if (is.null(original)) {
  message("Skipping categorical comparison plots until params$original_path exists.")
} else {
  categorical_cols <- {%categorical_cols}
  categorical_cols <- intersect(categorical_cols, shared)
  for (v in categorical_cols) {
    print(
      ggplot(combined, aes(x = .data[[v]], fill = .source)) +
        geom_bar(position = "dodge") +
        scale_fill_manual(values = dg_cols) +
        labs(title = v, x = v, fill = NULL) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 30, hjust = 1))
    )
  }
}
```

## DataGangeR comparison metrics

```{r dg-compare}
if (is.null(original)) {
  message("Skipping dataganger::compare_synthetic() until params$original_path exists.")
} else {
  comparison <- dataganger::compare_synthetic(original, synthetic)
  print(comparison)
}
```
