---
title: "DataGangeR: original vs synthetic comparison"
format:
  html:
    toc: true
    df-print: paged
params:
  original_path: "original_data.csv"
  synthetic_path: "synthetic_data.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 4
)
library(ggplot2)
```

{%purpose_line}

This report reads your original and synthetic datasets and compares them:
summary statistics, distribution plots (green = original, magenta = synthetic),
and DataGangeR's fidelity metrics. The **R** chunks run when you render this
document. The **Python** chunks are included for reference (they are not run):
copy them into your own Python environment, or set `eval: true` on them and
render with a configured Python toolchain.

Set the two file paths in the `params` block above (or edit the read chunks),
then render with Quarto.

## Read both datasets

```{r read}
original  <- utils::read.csv(params$original_path, stringsAsFactors = TRUE)
synthetic <- utils::read.csv(params$synthetic_path, stringsAsFactors = TRUE)

# Compare only the columns the two datasets share.
shared    <- intersect(names(original), names(synthetic))
original  <- original[, shared, drop = FALSE]
synthetic <- synthetic[, shared, drop = FALSE]

combined <- rbind(
  cbind(original,  .source = "original"),
  cbind(synthetic, .source = "synthetic")
)
combined$.source <- factor(combined$.source, levels = c("original", "synthetic"))

# DataGangeR convention: green = original, magenta = synthetic.
dg_cols <- c(original = "#3F8F4F", synthetic = "#D81B8C")
```

```{python}
#| eval: false
# Python reference (not executed)
import pandas as pd

original  = pd.read_csv("original_data.csv")
synthetic = pd.read_csv("synthetic_data.csv")

shared    = [c for c in original.columns if c in synthetic.columns]
original, synthetic = original[shared], synthetic[shared]
colors = {"original": "#3F8F4F", "synthetic": "#D81B8C"}
```

The original has `r nrow(original)` rows and the synthetic has
`r nrow(synthetic)` rows across `r length(shared)` shared columns.

## Summary statistics

```{r summary}
summary(original)
summary(synthetic)
```

```{python}
#| eval: false
original.describe(include="all")
synthetic.describe(include="all")
```

## Distribution comparisons

Green is the original data; magenta is the synthetic data.

```{r numeric-plots}
numeric_cols <- {%numeric_cols}
numeric_cols <- intersect(numeric_cols, shared)
for (v in numeric_cols) {
  print(
    ggplot(combined, aes(x = .data[[v]], fill = .source)) +
      geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
      scale_fill_manual(values = dg_cols) +
      labs(title = v, x = v, fill = NULL) +
      theme_minimal()
  )
}
```

```{r categorical-plots}
categorical_cols <- {%categorical_cols}
categorical_cols <- intersect(categorical_cols, shared)
for (v in categorical_cols) {
  print(
    ggplot(combined, aes(x = .data[[v]], fill = .source)) +
      geom_bar(position = "dodge") +
      scale_fill_manual(values = dg_cols) +
      labs(title = v, x = v, fill = NULL) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 30, hjust = 1))
  )
}
```

```{python}
#| eval: false
# Python reference (not executed)
import matplotlib.pyplot as plt

numeric_cols     = {%numeric_cols_py}
categorical_cols = {%categorical_cols_py}

for v in [c for c in numeric_cols if c in shared]:
    fig, ax = plt.subplots()
    for src, df in (("original", original), ("synthetic", synthetic)):
        ax.hist(df[v].dropna(), bins=30, alpha=0.5, label=src, color=colors[src])
    ax.set_title(v); ax.legend(); plt.show()

for v in [c for c in categorical_cols if c in shared]:
    cmp = pd.DataFrame({
        "original":  original[v].value_counts(normalize=True),
        "synthetic": synthetic[v].value_counts(normalize=True),
    }).fillna(0)
    cmp.plot.bar(color=[colors["original"], colors["synthetic"]], title=v)
    plt.show()
```

## DataGangeR comparison metrics

If the dataganger package is installed, this prints its built-in comparison:
per-variable standardized differences for numeric columns and total variation
distance for categorical columns.

```{r dg-compare}
if (requireNamespace("dataganger", quietly = TRUE)) {
  comparison <- dataganger::compare_synthetic(original, synthetic)
  print(comparison)
} else {
  message("Install the dataganger package for the built-in comparison metrics.")
}
```

```{python}
#| eval: false
# Python reference (not executed): simple standardized-difference drift check
for v in [c for c in numeric_cols if c in shared]:
    om, sm = original[v].mean(), synthetic[v].mean()
    osd = original[v].std() or 1.0
    print(f"{v}: standardized mean diff = {abs(om - sm) / osd:.3f}")
```
